Dissecting the causal question underlying the association between cancer and ADRD 

### 2021-04-12 04_reports/manuscript.Rmd
New version looking goood....

---
```{r, echo =  FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.align="center")
library(survival)
library(tidyverse)
library(tableone)
source(here::here("02_R", "03_descriptives.R"))
```

## Introduction

@@

## Overview of the causal structure and assumptions related to cancer diagnosis as a proxy for Pin1

@@

## Overview of the causal question and assumptions related to death as a competing event of ADRD

@@

## Motivating example with the RS data

In this section we illustrate how previous DAGS may reflect on the analysis performed to unveil the effect of Pin1 in all cause-dementia. We use data collected in the Rotterdam Study, a population-based prospective cohort study among persons living in the Ommoord district in Rotterdam, the Netherlands. Recruitment and initial assessments were held between 1990 and 1993, participants had follow-up visits between 1993-1995, 1997-1999, 2002-2005, and 2008-2010. All participants had data on incident cancer diagnosis and incident dementia diagnosis through follow-up, collected from medical records of general practitioners (including hospital discharge letters) and through linkage with national registries. Date and cause of death was collected on a weekly basis via municipal population registries. Data on clinical outcomes was available until 2015.

To match the analysis to our initial question, we considered as inclusion criteria being between 60 and 70 years old at study entry, without history of cancer diagnosis, free of cognitive decline or with previous history of dementia. Out of 7983 persons who participated at baseline, `r total_n` were considered eligible. Time to cancer diagnosis, to dementia diagnosis and death status was measured for all participants. All participants were followed from study entry until dementia diagnosis, death or 20 years after their individual baseline date, whichever occurred first.

@@

We present the following analysis:

The difference in dementia risk treating cancer as ever vs. never (As if Pin1 and cancer dx. happened at the same time if looked retrospectively).

We begin illustrating the most simple scenario (as Figure 1), were Pin1 overexpression is defined as "ever vs. never" and we consider death as an independent censoring mechanism. We consider baseline covariates for adjustment such as age at study entry, sex, educational attainment, smoking status, alcohol intake status.

## Discussion

@@

## Table 1

@@

>>>>

---
```{r, echo =  FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.align="center")
library(survival)
library(tidyverse)
library(tableone)
source(here::here("02_R", "03_descriptives.R"))
source(here::here("02_R", "05_plots.R"))
source(here::here("02_R", "05b_plot_causes_death.R"))
```

## 1. Introduction

@@

## 2. Overview of the causal structure and assumptions related to cancer diagnosis as a proxy for Pin1

@@

## 3. Overview of the causal question and assumptions related to death as a competing event of ADRD

@@

## 4. Motivating example with the RS data

In this section we illustrate how previous DAGS may reflect on the analysis performed to unveil the effect of Pin1 in all cause-dementia. We use data collected in the Rotterdam Study, a population-based prospective cohort study among persons living in the Ommoord district in Rotterdam, the Netherlands. Recruitment and initial assessments were held between 1990 and 1993; it was later extended between 2000 and 2001 consisting of individuals who had reached the age of 55 years or who had moved into the study area. Participants from first subcohort had follow-up visits between 1993-1995, 1997-1999, 2002-2005, and 2008-2010, second subcohort had follow-up visits between 2004 and 2005, and between 2011 and 2012. All participants had data on incident cancer diagnosis and incident dementia diagnosis through follow-up, collected from medical records of general practitioners (including hospital discharge letters) and through linkage with national registries. Date and cause of death was collected on a weekly basis via municipal population registries. Data on clinical outcomes was available until 2015.

To match the analysis to our initial question, we considered as inclusion criteria being between 60 and 70 years old at study entry, without history of cancer diagnosis, free of cognitive decline or with previous history of dementia. Out of 10998 persons who participated at study entry, `r total_n` were considered eligible. Time to cancer diagnosis, time to dementia diagnosis and death status was measured for all participants. All participants were followed from study entry until dementia diagnosis, death or 20 years after their individual baseline date, whichever occurred first.

### 4.1. Methods

First we illustrate the most simple scenario (as Figure 1), were Pin1 over-expression is defined as "cancer ever vs. never" as if Pin1 and cancer diagnosis happened at the same time if looked retrospectively. Second, we consider cancer as a time-varying exposure independent of death. This can be represented as follows, at five years from study entry the observed unadjusted risk of cancer was `r cancer_5y` and the risk of death was `r death_5y` for the entire cohort, thus we assume that those who died would have the same risk of cancer had they not died, unconditionally, over follow-up. Third, we consider time to cancer as the

To address confounding we fit inverse probability treatment weights including age at study entry, sex, educational attainment, smoking status and cohort as covariates.

To identify the controlled direct effect we considered death as an independent censoring mechanism conditional on time-varying shared causes of death and dementia.

### 4.2. Results

@@

## 5. Discussion

@@

\newpage

## Tables

### Table 1: Descriptive characteristics of individuals who had a cancer diagnosis and of those free of cancer diagnosis over follow up.

@@

\newpage

### Table 2. Risk difference and risk ratio for the risk of dementia

\newpage

## Figures

### Figure 1: Distribution of participants under each health status, by age over follow-up

```{r}
plot_distribution
```

### Figure 2. Causes of death for participants with and without incident cancer

```{r}
causes_death
```
